An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-14 article/latex_files/ch2/2-4-ELFI.tex
reordered repo

\pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package \pkg{NetworkX},
which supports general-purpose graphs. In most cases, the structure of
a likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
parameters of interest, i.e. those we try to infer a posterior
distribution, as the ones defined through the \code{elfi.Prior}
class. Finally, the observations are passed to the appropriate node
through the argument
\code{observed}.

>>>>

\pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package
\pkg{NetworkX}~\cite{hagberg2008exploring}, which supports
general-purpose graphs. In most cases, the structure of a
likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
parameters of interest, i.e. those we try to infer a posterior
distribution, as the ones defined through the \code{elfi.Prior}
class. Finally, the observations are passed to the appropriate node
through the argument \code{observed}.

### 2022-02-14 article/latex_files/ch3/3-0-design-principles.tex
reordered repo

\begin{algorithm}[!ht]
	\caption{ROMC. Requires the prior \( p(\thetab) \), the simulator
    \(M_r(\thetab)\), number of optimisation problems \(n_1\), number
    of samples per region \(n_2\), acceptance limit
    \(\epsilon\)}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{}
    \For{\(i \gets 1 \textrm{ to } n_1\)}
    \State \(\vb_i \sim p(\vb)\) \Comment{Draw nuisance variables}
    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \vb=\vb_i) \) \Comment{Define deterministic simulator}
      \State \( d_i(\thetab) = ||g(\thetab, \vb=\vb_i) - \yb_0||_2 \) \Comment{Define distance function}
      \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimisation problem}
      \If{\(d^*_i > \epsilon\)}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Estimate \(\mathcal{\hat{S}}_i\) and define \(q_i\) \Comment{Estimate proposal area}
      \State (Optional) Fit \(\Tilde{d}_i\) on \(\mathcal{\hat{S}}_i\) \Comment{Fit surrogate model}
      \\\hrulefill
      \For{\(j \gets 1 \textrm{ to } n_2\)}
      \State \(\thetab_{ij} \sim q_i\), compute \(w_{ij}\) as in Algorithm \ref{alg:sampling_GB} \Comment{Sample}
      \EndFor
    \EndFor
    \State \(E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
    \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalised posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
\( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\), (b) computing an
expectation \(E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
unnormalised posterior \(p_{d, \epsilon}(\thetab|\data)\). Sampling is
performed by getting \(n_2\) samples from each proposal distribution
\(q_i\). For each sample \(\thetab_{ij}\), the distance
function\footnote{As before, if a surrogate model \(\hat{d}\) is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing the expectation is easy after
weighted samples are obtained using the equation~\ref{eq:expectation},
so we do not discuss it separately. Evaluating the unnormalised
posterior requires the deterministic functions \(g_i\) and the prior
distribution \(p(\thetab)\)\footnote{There is no need for solving the
  optimisation problems and building the proposal regions.}. The
evaluation requires iterating over all \(g_i\) and evaluating the
distance from the observed data.

\begin{algorithm}[!ht]
	\caption{Approximation of \(\mathcal{\hat{S}}_i\) with a bounding box; 
  Requires: a model of distance \(d_i(\thetab)\), 
  an optimal point \(\thetab_i^*\), 
  a number of refinements \(K\), 
  a step size \(\eta\_\text{start}\), 
  maximum iterations \(M\) and 
  a curvature matrix \(\hessian_i\) (\(\jac_i^T\jac_i \) or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors \(\vb_{d}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,||\thetab ||)\)}
	\For{\(d \gets 1 \textrm{ to } D\)}
		\State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
		\State \(k \gets 0\)
		\State \(\eta \gets \eta\_\text{start}\) \Comment{Initialize \(\eta\)}
		\Repeat
          \State \(j \gets 0\)
        	\Repeat
            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}\) \Comment{Large step size \(\eta\).}
            \State \(j \gets j + 1\)
        	\Until{\(d(f_i(\Tilde{\thetab}), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}\)
        	\State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
        	\State \(k \gets k + 1\)
      \Until \(k = K\)
      \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d\) \Comment{Then, make the minimum step}
      \EndIf \label{algstep:box_constr_end}
    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
	\EndFor
	\State Fit a rectangular box around the region end points and define \(q_i\) as uniform distribution
	\end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}[!ht]
	\caption{ROMC. Requires the prior \( p(\thetab) \), the simulator
    \(M_r(\thetab)\), number of optimisation problems \(n_1\), number
    of samples per region \(n_2\), acceptance limit
    \(\epsilon\)}\label{alg:romc_algorithm}
	\begin{algorithmic}[1]
    \Procedure{ROMC}{}
    \For{\(i \gets 1 \textrm{ to } n_1\)}
    \State \(\vb_i \sim p(\vb)\) \Comment{Draw nuisance variables}
    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \vb=\vb_i) \) \Comment{Define deterministic simulator}
      \State \( d_i(\thetab) = ||g(\thetab, \vb=\vb_i) - \yb_0||_2 \) \Comment{Define distance function}
      \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimisation problem}
      \If{\(d^*_i > \epsilon\)}
        \State Go to 2 \Comment{Filter solution}
      \EndIf
      \State Estimate \(\mathcal{\hat{S}}_i\) and define \(q_i\) \Comment{Estimate proposal area}
      \State (Optional) Fit \(\Tilde{d}_i\) on \(\mathcal{\hat{S}}_i\) \Comment{Fit surrogate model}
      \\\hrulefill
      \For{\(j \gets 1 \textrm{ to } n_2\)}
      \State \(\thetab_{ij} \sim q_i\), compute \(w_{ij}\) as in Algorithm \ref{alg:sampling_GB} \Comment{Sample}
      \EndFor
    \EndFor
    \State \(\E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
    \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalised posterior}
    \EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
\( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\), (b) computing an
expectation \(\E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
unnormalised posterior \(p_{d, \epsilon}(\thetab|\data)\). Sampling is
performed by getting \(n_2\) samples from each proposal distribution
\(q_i\). For each sample \(\thetab_{ij}\), the distance
function\footnote{As before, if a surrogate model \(\hat{d}\) is
  available, it can be utilised as the distance function.} is
evaluated for checking if it lies inside the acceptance
region. Algorithm~\ref{alg:sampling_GB} defines the steps for
computing a weighted sample. Computing the expectation is easy after
weighted samples are obtained using the equation~\ref{eq:expectation},
so we do not discuss it separately. Evaluating the unnormalised
posterior requires the deterministic functions \(g_i\) and the prior
distribution \(p(\thetab)\)\footnote{There is no need for solving the
  optimisation problems and building the proposal regions.}. The
evaluation requires iterating over all \(g_i\) and evaluating the
distance from the observed data.

\begin{algorithm}[!ht]
	\caption{Approximation of \(\mathcal{\hat{S}}_i\) with a bounding box; 
  Requires: a model of distance \(d_i(\thetab)\), 
  an optimal point \(\thetab_i^*\), 
  a number of refinements \(K\), 
  a step size \(\eta\_\text{start}\), 
  maximum iterations \(M\) and 
  a curvature matrix \(\hessian_i\) (\(\jac_i^T\jac_i \) or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors \(\vb_{d}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,||\thetab ||)\)}
	\For{\(d \gets 1 \textrm{ to } D\)}
		\State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
		\State \(k \gets 0\)
		\State \(\eta \gets \eta\_\text{start}\) \Comment{Initialize \(\eta\)}
		\Repeat
          \State \(j \gets 0\)
        	\Repeat
            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}\) \Comment{Large step size \(\eta\).}
            \State \(j \gets j + 1\)
        	\Until{\(d(f_i(\Tilde{\thetab}), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}\)
        	\State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
        	\State \(k \gets k + 1\)
      \Until \(k = K\)
      \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d\) \Comment{Then, make the minimum step}
      \EndIf \label{algstep:box_constr_end}
    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
    	\State Runc steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
	\EndFor
	\State Fit a rectangular box around the region end points and define \(q_i\) as uniform distribution
	\end{algorithmic}
\end{algorithm}

### 2022-02-14 article/latex_files/ch4/4-2-inference.tex
reordered repo

\noindent
This is the basic inference utility of the ROMC implementation; we
draw $n_2$ samples for each bounding box region. This gives a total of
$k \times n_2$, where $k < n_1$ is the number of the optimal points
remained after filtering\footnote{From the $n_1$ optimisation
  problems, only the ones with $d_i(\thetab_*) < \epsilon$ are
  maintained for building a bounding box}. The samples are drawn from
a uniform distribution $q_i$ defined over the corresponding bounding
box and the weight $w_i$ is computed as in
Algorithm~\ref{alg:sampling_GB}.

@@

\noindent
This function computes the expectation
$E_{p(\thetab|\data)}[h(\thetab)]$ using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
built-in tools to get a summary of the obtained samples. In Figure
\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside the samples obtained from it. Finally, in the
code snippet, we demonstrate how to use the \code{compute_expectation}
function; in the current example, we define \code{h} to compute
firstly the empirical mean and afterwards the empirical variance. In
both cases, the empirical result is close to the ground truth
$\mu = 0$ and $\sigma^2 = 1$. The
\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
point $\theta$ using expression \eqref{eq:approx_posterior_omc}. The
\code{romc.eval_posterior(theta)} approximates the partition function
$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth posterior.

@@

% \begin{figure}[h]
%     \begin{center}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
%     \end{center}
%   \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples. (b) Right: Acceptance region around $\theta_1^*$ with the obtained samples plotted inside.}
%   \label{fig:example_sampling}
% \end{figure}

>>>>

\noindent
This is the basic inference utility of the ROMC implementation; we
draw \(n_2\) samples for each bounding box region. This gives a total of
\(k \times n_2\), where \(k < n_1\) is the number of the optimal points
remained after filtering\footnote{From the \(n_1\) optimisation
  problems, only the ones with \(d_i(\thetab_*) < \epsilon\) are
  maintained for building a bounding box}. The samples are drawn from
a uniform distribution \(q_i\) defined over the corresponding bounding
box and the weight \(w_i\) is computed as in
Algorithm~\ref{alg:sampling_GB}.

@@

\noindent
This function computes the expectation
\(\E_{p(\thetab|\data)}[h(\thetab)]\) using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
\(Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab\); with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
built-in tools to get a summary of the obtained samples. In Figure
\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside the samples obtained from it. Finally, in the
code snippet, we demonstrate how to use the \code{compute_expectation}
function; in the current example, we define \code{h} to compute
firstly the empirical mean and afterwards the empirical variance. In
both cases, the empirical result is close to the ground truth
\(\mu = 0\) and \(\sigma^2 = 1\). The
\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
point \(\theta\) using expression \eqref{eq:approx_posterior_omc}. The
\code{romc.eval_posterior(theta)} approximates the partition function
\(Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab\) using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth posterior.

@@

% \begin{figure}[h]
%     \begin{center}
%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
%     \end{center}
%   \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples. (b) Right: Acceptance region around \(\theta_1^*\) with the obtained samples plotted inside.}
%   \label{fig:example_sampling}
% \end{figure}

### 2022-02-14 article/latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex
reordered repo

title={ROMC (gradient-based) - \(\displaystyle \theta_2\)},

@@

\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00551481774368941);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0605814218347914);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.168699423699822);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.340262186728983);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.48172967019454);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.661723210849365);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.926167683238105);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.1131441320768);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.19038558853121);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.39560413532974);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.38890799766854);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.45486223069012);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.89014334843844);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.80399178478282);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.15561058973456);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.33061276076321);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24024442959031);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.34890403251458);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.31095726097247);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.33900171768764);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.07598627244674);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.71987564844275);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65165022964706);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.40256222683377);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18758980571923);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.0980208933351);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.803715750920935);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.742179116696122);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.678702875114039);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.583516431505742);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.398341719580722);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.257727424376927);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.229546935490747);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.210112717761985);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129179090828181);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0899034779939157);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0431810229330879);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0479789143700981);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0287873486220588);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0143936743110293);

@@

0.0318388478807404 0
0.0318388478807404 3

@@

\addlegendentry{$\mu = 0.032$}

@@

-0.177435597378694 1
0.197481062716323 1

@@

\addlegendentry{$\sigma = 0.170$}

>>>>

title={ROMC (Neural Network) - \(\displaystyle \theta_2\)},

@@

\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0687105672242381);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.184372208227465);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.251601377874948);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.39271865402813);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.516086691042045);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.829805712987072);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.847041859829425);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.888673668222991);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.16327691119626);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.37486835291373);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.19991232691807);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.42636185782617);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.39151225358654);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.82926351678107);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.94594508515498);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.2129470534662);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.34655785188779);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.63158719290653);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.55685248411093);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.63080691718184);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.52614922878044);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.93355077757299);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.72788273613444);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.36820910319437);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.0258751741395);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,0.915348210439271);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.784575688430834);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.649930571025344);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.528011081166735);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.423788895495386);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.255909163140911);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.28970067094874);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.19412252617457);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.189105171805318);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.127165545733375);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0974061925124961);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0718579504601718);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0457277866564734);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0587928685583229);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0979881142638707);

@@

0.0301562808023854 0
0.0301562808023854 3

@@

\addlegendentry{$\mu = 0.030$}

@@

-0.180267449762181 1
0.196611267527429 1

@@

\addlegendentry{$\sigma = 0.171$}

### 2022-02-14 latex/article.tex
reordered repo

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}

@@

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Vasilis Gkolemis\\ATHENA RC
  \And Michael Gutmann\\University of Edinburgh
   \And Henri Pesonen \\University of Oslo
}
\Plainauthor{Vasilis Gkolemis, Michael Gutmann}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}

%% - \Abstract{} almost as usual
\Abstract{ Performing the inference in statistical models with
  intractable likelihood is challenging; most likelihood-free
  inference (LFI) methods encounter problems in accuracy and
  efficiency. In this paper, we present the implementation of Robust
  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
  LFI framework that provides accurate weighted samples from the
  posterior. Our implementation can be used in two ways. Firstly, a
  scientist may use it as an out-of-the-box LFI algorithm; we provide
  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
  framework.  Therefore, it is straightforward to compare it with the
  rest of the methods included in \pkg{ELFI}. Afterwards, we have
  carefully split ROMC into isolated algorithmic steps. Hence, we
  allow the researcher to experiment by replacing part(s) of the
  algorithm with novel subcomponent(s). In both scenarios, all steps
  of ROMC run in a fully-parallelizable manner, exploiting all CPU
  cores. We also provide helpful functionalities for (i) inspecting
  the inference process and (ii) evaluating the obtained
  samples. Finally, we test the robustness of our implementation on
  some typical LFI examples.  }

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Bayesian Inference, Simulator-based models, Likelihood-free, \proglang{Python}, \proglang{ELFI}}
\Plainkeywords{Bayesian Inference, Simulator-based models, Likelihood-free, Python, ELFI}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Vasilis Gkolemis\\
  Information Management Systems Institute (IMSI)\\
  ATHENA Research and Innovation Center\\
  Athens, Greece\\
  E-mail: \email{vgkolemis@athenarc.gr}\\
  URL: \url{https://givasile.github.io}
}

@@

%% See jss_template_article.tex for formatting instructions and macros to e.g. typset code
%% I only kept the informative sidebars as comments

@@

@@

\subsection{Engine for likelihood-free inference (ELFI)}
\label{subsec:ELFI}
\input{latex_files/ch2/2-4-ELFI.tex}

>>>>

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{orcidlink}
\usepackage{amsmath,amssymb,bbold, xfrac}

@@

\author{Vasilis Gkolemis~\orcidlink{0000-0002-2636-0245}\\ATHENA RC \And
  Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
}
\Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}

\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable ROMC implementation}

%% - \Abstract{} almost as usual
\Abstract{ Performing the inference in statistical models with
  intractable likelihood is challenging; most likelihood-free
  inference (LFI) methods encounter problems in accuracy and
  efficiency. In this paper, we present the implementation of Robust
  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
  LFI framework that provides accurate weighted samples from the
  posterior. Our implementation can be used in two ways. Firstly, a
  scientist may use it as an out-of-the-box LFI algorithm; we provide
  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
  framework.  Therefore, it is straightforward to compare it with the
  rest of the methods included in \pkg{ELFI}. Afterwards, we have
  carefully split ROMC into isolated algorithmic steps. Hence, we
  allow the researcher to experiment by replacing part(s) of the
  algorithm with novel subcomponent(s). In both scenarios, all steps
  of ROMC run in a fully-parallelizable manner, exploiting all CPU
  cores. We also provide helpful functionalities for (i) inspecting
  the inference process and (ii) evaluating the obtained
  samples. Finally, we test the robustness of our implementation on
  some typical LFI examples.}

%% - \Keywords{}
\Keywords{Bayesian inference, implicit models, likelihood-free, \proglang{Python}, \pkg{ELFI}}
\Plainkeywords{Bayesian inference, implicit models, likelihood-free, Python, ELFI}

%% - \Address{} of at least one author
\Address{
  Vasilis Gkolemis\\
  Information Management Systems Institute (IMSI)\\
  ATHENA Research and Innovation Center\\
  Athens, Greece\\
  E-mail: \email{vgkolemis@athenarc.gr}\\
  URL: \url{https://givasile.github.io}
}

@@

@@

To the best of our knowledge, this is the first attempt to implement
the ROMC inference method to a generic LFI framework. Therefore, there
are not other implementation directly comparable to ours. For this
reason, we test the implementation against (i) an artificial example
with tractable likelihood and (ii) the second-order moving average
(MA2) example from the \pkg{ELFI} package. In the latter, we generate
data using known parameters and we compare our results with the
typical Rejection ABC approach.

@@

\subsection[Engine for likelihood-free inference (ELFI)]{\pkg{Engine for likelihood-free inference (ELFI)}}
\label{subsec:ELFI}
\input{latex_files/ch2/2-4-ELFI.tex}
